---
title: "Profile of preference: Rashtree analysis"
author: "Benini & Sons"
date: "2019-12-10"
categories:
  - Indicator
  - Prioritisation
tags:
  - Okular-Analytics
  - Benini
  - UNHCR
  - Edouard-Legoupil
---


Often affected population or key informant are requested to share their preference for specific type of interventions.  How can we identify patterns of preference within a dataset? Can we identify  groupings  on  the  basis of several (categorical or continuous) variables that differentiate profiles optimally.

<!--MORE-->

This tutorial is based on the publication [Priorities and preferences in humanitarian needs assessments -Their measurement by rating and ranking methods](http://aldo-benini.org/Level2/HumanitData/OkularAnalytics_Priorities_and_Preferences_v2019_01_23.pdf).

## Loading packages

```{r setup, include = TRUE, message = FALSE, warning = FALSE, echo = TRUE}
## This function will retrieve the packae if they are not yet installed.
using <- function(...) {
   libs <- unlist(list(...))
    req <- unlist(lapply(libs,require,character.only = TRUE))
    need <- libs[req == FALSE]
    if (length(need) > 0) { 
        install.packages(need)
        lapply(need,require,character.only = TRUE)
    }
}

## Getting all necessary package
using("foreign", "PlackettLuce", "tidyverse", "qvcalc","kableExtra","stargazer","NLP",
                     "ggthemes", "ggrepel", "GGally", "bbplot","ggpubr",'grid','gridExtra', 'forcat', 'psychotree')

rm(using)

# This small function is used to have nicely left align text within charts produced with ggplot2
left_align <- function(plot_name, pieces){
  grob <- ggplot2::ggplotGrob(plot_name)
  n <- length(pieces)
  grob$layout$l[grob$layout$name %in% pieces] <- 2
  return(grob)
}


```

## Dataset

The data used in the this tutorial comes from NPM Site Assessment Round 11, 2018 - for  the  Rohingya  refugee camps in Bangladesh.

```{r , include=TRUE, echo = TRUE}
data_csv <- read.csv(file = "181227_2027AB_NPM11_Priorities_PlackettLuce.csv", header = T, sep = ",")


 
used_var <- as.character(names(data_csv))[grepl("l_",as.character(names(data_csv)))]

# Structure of the data:
#str(data_csv)
```


## Can we identify differences within preferences?


In the dataset, there are variables define groups among which sectoral needs priorities may differ significantly.This is corresponds to the detection of __Differential Item Functioning__ (DIF).


For instance, we have 4 sub-districts with refugee settlements (`upazila`), a continuous population size variable for the 1,990 camp blocks (`log10pop`) (logarithmic), and the distance from the nearest health care facility (`healthWalk_enc`, with five levels) as a marginalization indicator. These 

[Rashtree Visualisation](https://cran.r-project.org/web/packages/psychotree/vignettes/raschtree.pdf) are designed to identify significant differences within preferences.

We first need to format the data so that it can be consumed by the algorithm.

```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 10, fig.height=12, size="small"}
covariate <- data_csv[ , c("upazila","log10pop","healthWalk_enc")]

resp <- as.matrix(data_csv[ , used_var])

## Rashtree accepts only 0 or 1 - so everything above 0 shalle be replaced by 1
for (i in 1:nrow(resp)) {
  for (j in 1:ncol(resp)) if (resp[i,j] > 0 ) resp[i,j] = 1
}
## resp will be a matrix variable used in the model
covariate$resp <- resp

# To exclude rows where all observed item responses are either 0 or 1, 
# we select only the subsetof cases for which the proportion of correct item responses is strictly
# between 0 and 1 forfuther analysis.
covariate <- subset(covariate, rowMeans(resp, na.rm = TRUE) > 0 &   rowMeans(resp, na.rm = TRUE) < 1)
``` 

We can now compute and display it. 

If the Rasch tree shows at least one split, DIF is present and there are groups with significant difference in their hiearchy of needs.


```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 15, fig.height=12, size="small"}
## Compute the rashtree
raschtree <- raschtree(resp ~ upazila + log10pop + healthWalk_enc, data = covariate)

## and plotting it...
plot(raschtree)
```


We can now extract the groups and descriptions.

```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 10, fig.height=12, size="small"}
## Compute the rashtree

kable( as.data.frame(itempar(raschtree)))


```

We can also explore trees by focusing on one aspect...


```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 15, fig.height=12, size="small"}

## and plotting it...
plot(raschtree(resp ~ upazila, data = covariate),
     main = "For upazila")
```



```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 15, fig.height=12, size="small"}

## and plotting it...
plot(raschtree(resp ~ log10pop , data = covariate),
     main = "For population")
```



```{r , echo = TRUE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = "", fig.width = 15, fig.height=12, size="small"}

## and plotting it...
plot(raschtree(resp ~  healthWalk_enc, data = covariate),
     main = "For distance to Health Center")
```




