---
title: "rhdx: a R package to interact with HDX"
author: Ahmadou Dicko
date: '2018-12-23'
categories:
  - Data
tags:
  - HDX
  - WebData
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, warning = FALSE, message = FALSE)
```

rhdx is an R package to interact with the [HDX](https://data.humdata.org/) API.

It provide a series of utilities to facilitates interaction & analysis.

The Humanitarian Data Exchange (HDX) is an open platform for sharing data across crises and organisations. Launched in July 2014, the goal of HDX is to make humanitarian data easy to find and use for analysis. Our growing collection of datasets has been accessed by users in over 200 countries and territories.

<!--MORE-->



# Install `rhdx`


```{r, eval = TRUE}
remotes::install_gitlab("dickoa/rhdx")
```

```{r, eval = FALSE}
remotes::install_github("dickoa/rhdx")
remotes::install_gitlab("dickoa/rhdx")
```

# Connect to the HDX server

```{r, cache = FALSE}
library(rhdx)
set_rhdx_config(hdx_site = "prod")
```

You can check your config

```{r}
get_rhdx_config()
```

# Search datasets

## List datasets matching a pattern
```{r}
library(tidyverse)
search_datasets("ACLED Mali", rows = 2)
```

## Select the dataset

We will select the first dataset

```{r}
search_datasets("ACLED Mali", rows = 2) %>%
  nth(1)
```

## List dataset resources

```{r}
search_datasets("ACLED Mali", rows = 2) %>%
  nth(1) %>%
  get_resources()
```

## We select the first resource and read it directly in R

```{r, cache = TRUE}
search_datasets("ACLED Mali", rows = 2) %>%
  nth(1) %>%
  get_resource(1) %>%
  read_resource(filename = "acled_mali.csv", hxl = TRUE)
```

## Save it for further use

```{r}
search_datasets("ACLED Mali", rows = 2) %>%
  nth(1) %>%
  get_resource(1) %>%
  read_resource(filename = "acled_mali.csv", hxl = TRUE) -> acled_mali
```

# Now let's see where we have these conflicts in Mali

## First step: just select the coordinates

```{r}
acled_mali %>%
  select(longitude, latitude) %>%
  distinct() -> acled_mali_coord
acled_mali_coord
```

## Second step download the official Mali COD boundaries from HDX

```{r}
pull_dataset("administrative-boundaries-cod-mli") %>%
  get_resources()
```

We just need the country boundary (admin 0) which is the fourth resource

```{r}
pull_dataset("administrative-boundaries-cod-mli") %>%
  get_resource(4) %>%
  read_resource() -> adm0_mali
```

`ggplot2` is a R package for graphics with a nice grammar to compose any type of visual, we also use `sf` which add spatial data support.

```{r}
library(sf)
ggplot() +
  geom_sf(data = adm0_mali) +
  geom_point(data = acled_mali_coord, aes(x = longitude, y = latitude)) +
  theme_bw()
```
