---
title: "Fatal events in the Sahel"
author: "Ahmadou Dicko"
date: "2019-02-27"
categories:
  - DataViz
  - Animation
  - rhdx
  - API
tags:
  - Ahmadou-Dicko
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this short post (crossposted [here](https://www.ahmadoudicko.com/post/sahel-fatalities/)), we will show how to use the `rhdx`, `dplyr`, `purrr`, `sf` and `gganimate` R packages to show the number of fatal incidents in 5 Sahelian countries.

<!--MORE-->

The [`rhdx`](https://gitlab.com/dickoa/rhdx) package is not yet on CRAN, so you will need to use the [`remotes`](https://cran.r-project.org/web/packages/remotes/index.html) package to install it first:

```{r, eval=FALSE}
remotes::install_gitlab("dickoa/rhxl") ## rhdx dependency
remotes::install_gitlab("dickoa/rhdx") ## github mirror also avalailable

install.packages("gifski")

```

This analysis was inspired by this tweet by José Luengo-Cabrera, researcher at the Crisis Group.

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">G5 Sahel: conflict-related fatalities totaled 2,832 in 2018, a 74% increase relative to 2017.<br><br>- On average, 63% of fatalities have been concentrated in Mali since 2012.<br>- Last year, fatalities were largely located in central Mali, the Liptako-Gourma region &amp; the Lake Chad basin. <a href="https://t.co/feRtcxsScb">pic.twitter.com/feRtcxsScb</a></p>&mdash; José Luengo-Cabrera (@J_LuengoCabrera) <a href="https://twitter.com/J_LuengoCabrera/status/1100340244535263232?ref_src=twsrc%5Etfw">February 26, 2019</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

Our visualization will be done for the following countries in the Sahel : _Burkina Faso_, _Mali_, _Chad_, _Mauritania_ and _Niger_.
There is a lot of insecurity and conflicts in these countries that resulted in the death of several thousands of people.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(rhdx)
library(gganimate)
```

The goal of this post is to visualize the number of fatal events in theses countries between 2012 and 2018 using an animated map. 
In order to do that, will need to get the administrative boundaries for these countries. 
We can get the latest boundaries validated by the governments and the humanitarian community directly from [_HDX_](https://data.humdata.org/) using the `rhdx` package.
We will use `rhdx::pull_dataset` function and use the name of the dataset of interest as value. `rhdx::get_resource` and `rhdx::download_resource` allow to respectively get the a resource
by its index and read the data into memory.

```{r}
set_rhdx_config()

wca <- pull_dataset("west-and-central-africa-administrative-boundaries-levels") %>%
  get_resource(1) %>%
  read_resource()
glimpse(wca)
```

`wca` is a [`Simple Feature`](https://r-spatial.github.io/sf/articles/sf1.html) and we can manipulate it using `sf` and `dplyr`. 
The data downloaded from _HDX_ covers the 24 countries of West and Central Africa, we will filter the data to extract the 5 countries of interest.

```{r}
g5_ab <- wca %>%
  filter(admin0Pcod %in% c("BF", "ML", "NE", "MR", "TD"))
```
We can now check our data by plotting it

```{r}
g5_ab %>%
  ggplot() +
  geom_sf() +
  theme_minimal()
```

Now that we have our background map, the next step is to get the _conflict data_. One of the main source for conflict data in the Sahel is [ACLED](https://www.acleddata.com/)
and it can also be accessed from _HDX_. We will use the `rhdx::search_datasets` function since it allows to access multiple datasets as `list`. 
_HDX_ is based on CKAN whose search is powered by [`SOLR`](http://lucene.apache.org/solr/). SOLR can be used to build complex queries to get exactly the datasets we want. 
In our case, we want conflict data from the [ACLED organization](http://data.humdata.org/organization/acled) in _HDX_ and from the 5 countries (`group` in CKAN API)

```{r, message=FALSE, warning=FALSE}
solr_query <- "organization:acled AND groups:(mli OR bfa OR tcd OR mrt OR ner)"
g5_acled <- search_datasets(query = "conflict data",
                            fq = solr_query)
g5_acled
```

We will select the first 5 datasets (our 5 countries) in the list of datasets and bind them together using `purrr::map_df` and a helper function.

```{r, message=FALSE, warning=FALSE}
g5_acled <- g5_acled[1:5] ## pick the first 5 the 6th is the Africa wide dataset

# ## create a helper function to read resources from each dataset
# read_acled_data <- function(dataset) {
#   dataset %>%
#     get_resource(1) %>%
#     read_resource(force_download = TRUE)
# }
# 
# g5_acled_data <- map_df(g5_acled, read_acled_data)
# glimpse(g5_acled_data)
```

We have all the data we need for our analysis, we just need to aggregate total number fatalities by geographical coordinates, countries and year.

```{r, message=FALSE, warning=FALSE}
# g5_acled_fatalities_loc <- g5_acled_data %>%
#   filter(year %in% 2012:2018, fatalities > 0) %>%
#   group_by(year, country, latitude, longitude) %>%
#   summarise(total_fatalities = sum(fatalities, na.rm = TRUE)) %>%
#   arrange(year) %>%
#   ungroup()
```

We can finally use our boundaries (`g5_ab`), the conflict data (`g5_acled_fatalities_loc`) with `gganimate` to dynamically visualize the different fatal incidents in G5 Sahel countries.

```{r anim_map, message=FALSE, warning=FALSE}
# g5_ab %>%
#   ggplot() +
#   geom_sf(fill = "#383838", color = "gray") +
#   coord_sf(datum = NA) +
#   geom_point(data = g5_acled_fatalities_loc, aes(longitude, latitude, size = total_fatalities, fill = total_fatalities), shape  = 21, color = "transparent") +
#   scale_fill_viridis_c(option = "plasma") +
#   geom_sf_text(aes(label = admin0Name), color = "gray", fontface = "bold") +
#   labs(x = "",
#        y = "",
#        title = "Fatal events in the Sahel G5",
#        subtitle = "for the year {current_frame}",
#        caption = "source: ACLED") +
#   theme_void() +
#   theme(legend.position = "none") +
#   transition_manual(year, cumulative = TRUE) +
#   shadow_mark()

```



Session info for this analysis.

<details>
<summary>Session info</summary>

```{r}
devtools::session_info()
```
</details>

